### 2015-05-01 paper.tex
paper

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}

%% Paper title.

\title{TRIAD: Towards Real-time, Interactive Analysis of Data}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Remco Chang \and Danyel Fisher \and Jeffrey Heer \and Carlos Scheidegger}
\authorfooter{
%% insert punctuation at end of each item
\item
 Remco Chang is with Tufts University. email: remco@cs.tufts.edu
\item
 Danyel Fisher is with Microsoft Research. email: danyelf@microsoft.com
\item
 Jeffrey Heer is with the University of Washington. email: jheer@uw.edu
\item
 Carlos Scheidegger is with the University of Arizona. email: cscheid@cs.arizona.edu
}

%other entries to be set up for journal
\shortauthortitle{Workshop proposal: TRIAD}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{}

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{}

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.
\section{Goals}
%% the only exception to this rule is the \firstsection command
The goal of this workshop is to foster innovative work in the backends
that will support the next generation of interactive data analysis
tools. We envision such backends to require novel insights in database
technology, algorithm design, and information visualization techniques.

In the current design of database and analytics systems, interactive
data analysis concerns are usually an afterthought. Although recent
architectural changes like columnar databases have enabled faster
visualization systems, these sorts of innovations have been designed
less for visualization than for accelerating queries in general. In
this workshop, we will explore the consequences of pushing the
concerns of interactive data analysis deep into the ``computational
fabric'' of our systems: the back-end infrastructure of data storage,
organization, and retrieval. If we design with the knowledge that
every analysis is eventually read by a human, bound by perceptual
limitations and latency constraints, what are the consequences for
database design, or data analysis and visualization algorithms? Are
there fundamental tradeoffs? What can we give up?

There have been previous workshops in VIS that have investigated
issues of large data visualization, such as LDAV (and LDAV will be a
co-located symposium this year at VIS). Our focus is markedly
different from that of LDAV. LDAV focuses on simulation and imaging
data, while we believe current techniques in information visualization
and visual analytics are close to hitting a ceiling in terms of
performance. We propose this workshop as a venue to discuss these
limits, and fundamental new ways to attack them.

\section{Technical Scope}
The scope of this workshop will span, among possibly other fields,
information visualization, visual analytics, algorithm design,
database architecture, and statistics. Although this is clearly a
broad scope, the workshop will focus on the backends for interactive
tools. We have made significant progress in our understanding of
perceptual and human-interaction issues behind interactive
visualization design. However, much of this understanding has not, in
general, been translated into new designs for our backend
infrastructure. We think now is the right time to have a discussion of
backend design for interactive data analysis. Following the principle
of ``No One Size Fits All'' in designing databases [7], recent
publications by the organizers have shown that by taking advantage of
needs and limitations of interactive visualization systems, databases
can be optimized by making tradeoffs in novel and interesting
ways. For example, imMens [5], profiler [2], and nanocubes [4] have
shown that careful algorithm design can enable extremely fast
interactive visualizations of massive data. Scalr [3] has shown that
traditional query planners can fruitfully take advantage of perceptual
and resolution limits. ForeCache [6] utilizes a predictive prefetching
strategy by learning a user’s interaction patterns and prefetching
data from server to support real-time visual exploration. Lastly,
SampleAction [1] takes advantage of streaming databases and
incorporates HCI principles that users need immediate feedback, but
are willing to wait for high-fidelity results.

In addition to the work by the organizers, researchers in the database
community have also made significant stride in developing databases
specifically for interactive visualization, analysis, and exploration
of data. For example, systems like Tableau utilizes columnar databases
which provide speedups for typical visual analytics tasks that require
operations across columns (instead of the traditional rows). BlinkDB
[8] uses a sampling strategy to build a smaller, but representative
database from a larger one. Map-D [9] takes advantage of modern
graphics card capabilities and resulted in a specialized database that
runs entirely within (an array of) graphics cards. Finally, SciDB [10]
breaks away from the traditional relational data structure and is
designed to store and query multidimensional scientific data that are
stored in an array-based structure.

Given the increasing number of publications and backend systems that
can be tailored or applied to interactive visual exploration of large
data, we believe that it is very important for the InfoVis and the
VAST communities to begin having discussions and to develop a research
agenda that focuses on the backend technologies. In this proposed
workshop, we seek to invite members from the database community
(including authors of the aforementioned papers) to meet with the VIS
community and begin such dialogs. As many of these database
researchers are current collaborators of the organizers (see Section
XXX below for a list of intended invitees), we believe that the
organizing team is well suited to influence research not only in
information visualization, but in these other related fields as well.

\section{Format, Length, and Planned Activities}
We propose a half-day workshop that focuses on discussion between the
VIS and database researchers. The workshop will solicit submissions of
position papers as well as short research papers from both the VIS and
database communities. These papers will be reviewed by the organizing
team. The authors of the selected top papers will be asked to each
give a short presentation of their work (or position).

The goal of the presentation is to allow the attendees to get an
overview of the different perspectives and approaches of data storage,
organization, and retrieval techniques that can be tailored
specifically to support interactive data visualization and
exploration. We anticipate that these presentations will take up
approximately the first half of the workshop.

Following a short break, the second half of the workshop will be
oriented around open discussions and possibly a question-and-answer
session. As we anticipate that members of the two different
communities will not have a common language or viewpoint, we believe
that the open-ended format will facilitate discussions and general
knowledge sharing.

It is relevant to note that the organizing team is also currently
seeking a followup ``closed-door'' (invitation only) meeting that will
take place in the afternoon outside of the main VIS venue. The purpose
of this closed-door meeting is to allow for a more intimate discussion
between enthusiastic researchers who are actively seeking
collaborations across the aisle. Although it is not the goal of the
organizing team to be exclusive, it is our belief that a more focused
discussion will have a higher likelihood towards developing a cohesive
research agenda. Note that this is not an extraordinary proposal: such
a format has been successfully employed by other workshops in previous
years at VIS, most notably by the BELIV workshop~\cite{beliv}.

\section{List of Participants and Invitees}

As noted earlier, it is our hope that this workshop will include both
attendees from the VIS and database communities. However, since the
workshop will be located at VIS, we will specifically reach out to
members of the database community and invite them to attend the
workshop. We will seek a mix of researchers in different stages of
their careers, together with researchers from different companies in
the industry. If possible, we will also seek to identify a more senior
person from this invited list as the keynote speaker for the workshop.

Senior researchers:

\begin{itemize}
\item Sam Madden (MIT)
\item Joe Hellerstein (Berkeley, Trifacta)
\item Mike Stonebraker (MIT)
\item Ugur Cetintemel (Brown)
\item Stan Zdonik (Brown)
\end{itemize}

Early or mid-career researchers:
\begin{itemize}
\item Eugene Wu (Columbia)
\item Aaron Elmore (U. Chicago)
\item Joey Gonzalez (Berkeley AMPLab, CMU, GraphLab/Dato)
\item Bill Howe (UW)
\item Arnab Nandi (OSU)
\item Cecilia Aragon (UW; astronomy/scivis background)
\item Jenny Duggan (Northwestern)
\item Mike Cafarella (Michigan)
\end{itemize}

Junior Researchers: (Students, etc)
\begin{itemize}
\item Manasi Vartak (MIT, SeeDB)
\item Todd Mostak (MapD)
\item Leilani Battle (MIT, Tile Prefetching)
\item Dominik Moritz (UW)
\item Holger Pirk (MIT postdoc, databases)
\end{itemize}

Relevant companies we will reach out to:
\begin{itemize}
\item Tableau
\item Trifacta
\item Paradigm4
\item Graphistry
\end{itemize}

\section{Requested Facilities}

We should ask for a setup split halfway with seats for lectures and space for posters in the back. TBD.

\section{Intended Result and Impact of Workshop Results}

TBD.

\section{Related publications by Organizers}
\cite{Battle:2013:DRO, Lins:2013:NFR, Liu:2013:IRT, Fisher:2012:TMI, Kandel:2012:PIS}
[6] Battle, Chang, Stonebraker. Dynamic Prefetching of Data Tiles for Interactive Visualization. VLDB, 2015 (Conditionally Accepted).

\section{References}

\cite{Agarwal:2013:BQW}

[7] Stonebraker and Cetintemel. "One Size Fits All": An Idea Whose Time Has Come and Gone. ICDE '05
[9] Mostak. An overview of MapD. http://map-d.com/docs/mapdwhitepaper. 2014
[10] Cudre-Mauroux et al. SciDB: A Science-Oriented DBMS. VLDB 2009.

\section{Organizer Bios}

\paragraph*{Remco Chang} is an Assistant Professor in the Computer Science Department at Tufts University. He received his BS from Johns Hopkins University in 1997 in Computer Science and Economics, MSc from Brown University in 2000, and PhD in computer science from UNC Charlotte in 2009. Prior to his PhD, he worked for Boeing developing real-time flight tracking and visualization software, followed by a position at UNC Charlotte as a research scientist. His current research interests include visual analytics, information visualization, and HCI. His research has been funded by NSF, DHS, MIT Lincoln Lab, and Draper. He received the NSF CAREER Award in 2015.

\paragraph*{Danyel Fisher} is a Senior Researcher in information visualization and human-computer interaction at Microsoft Research's VIBE group. His research focuses on ways to help users interact with data more easily. His recent work has looked at ways to make big data analytics faster and more interactive. Danyel received his MS from UC Berkeley, and his PhD from UC Irvine. He tweets at @FisherDanyel.

\paragraph*{Jeffrey Heer} is an Associate Professor of Computer Science \& Engineering at the University of Washington, where he directs the Interactive Data Lab and conducts research on data visualization, human-computer interaction and social computing. The visualization tools developed by his lab (D3.js, Vega, Protovis, Prefuse) are used by researchers, companies and thousands of data enthusiasts around the world. His group's research papers have received awards at the premier venues in Human-Computer Interaction (ACM CHI, UIST, CSCW) and Information Visualization (IEEE InfoVis, VAST, EuroVis). Other awards include MIT Technology Review's TR35 (2009), a Sloan Foundation Research Fellowship (2012), and a Moore Foundation Data-Driven Discovery Investigator award (2014). Jeff holds BS, MS and PhD degrees in Computer Science from UC Berkeley, whom he then betrayed by joining the Stanford faculty (2009-2013). Jeff is also a co-founder of Trifacta, a provider of interactive tools for scalable data transformation.

\paragraph*{Carlos Scheidegger} is an Assistant Professor in the Department of Computer Science at the University of Arizona since 2014. Prior to joining the Arizona faculty, Carlos worked at AT\&T Research from 2009 to 2014, where he helped develop award-winning, open-source techniques for massive data sources (http://nanocubes.net). His current research interests are in the intersection of large-scale data analysis, information visualization, data management, and software infrastructure for scientific collaboration. His research has received multiple awards at top venues in data visualization (IEEE InfoVis, IEEE SciVis, EuroVis).

\bibliographystyle{abbrv}
\bibliography{paper}
\end{document}

### 2015-05-01 paper.tex
added multibib. Cleaned up the references

@@
 \usepackage{times}
+\usepackage{multibib}
+\newcites{orgref}{Publications by Organizers}
+\newcites{ref}{Other References}
 
@@
 backend design for interactive data analysis. Following the principle
-of ``No One Size Fits All'' in designing databases [7], recent
+of ``No One Size Fits All'' in designing databases~\citeref{stonebraker2005one}, recent
 publications by the organizers have shown that by taking advantage of
@@
 can be optimized by making tradeoffs in novel and interesting
-ways. For example, imMens [5], profiler [2], and nanocubes [4] have
+ways. For example, imMens~\citeorgref{Liu:2013:IRT}, profiler~\citeorgref{Kandel:2012:PIS}, and nanocubes~\citeorgref{Lins:2013:NFR} have
 shown that careful algorithm design can enable extremely fast
-interactive visualizations of massive data. Scalr [3] has shown that
+interactive visualizations of massive data. Scalr~\citeorgref{Battle:2013:DRO} has shown that
 traditional query planners can fruitfully take advantage of perceptual
-and resolution limits. ForeCache [6] utilizes a predictive prefetching
+and resolution limits. ForeCache~\citeorgref{Battle:2015:VLDB} utilizes a predictive prefetching
 strategy by learning a user’s interaction patterns and prefetching
 data from server to support real-time visual exploration. Lastly,
-SampleAction [1] takes advantage of streaming databases and
+SampleAction~\citeorgref{Fisher:2012:TMI} takes advantage of streaming databases and
 incorporates HCI principles that users need immediate feedback, but
@@
 which provide speedups for typical visual analytics tasks that require
-operations across columns (instead of the traditional rows). BlinkDB
-[8] uses a sampling strategy to build a smaller, but representative
-database from a larger one. Map-D [9] takes advantage of modern
+operations across columns (instead of the traditional rows). BlinkDB~\citeref{Agarwal:2013:BQW}
+uses a sampling strategy to build a smaller, but representative
+database from a larger one. Map-D~\citeref{mapd} takes advantage of modern
 graphics card capabilities and resulted in a specialized database that
-runs entirely within (an array of) graphics cards. Finally, SciDB [10]
+runs entirely within (an array of) graphics cards. Finally, SciDB~\citeref{cudre2009demonstration}
 breaks away from the traditional relational data structure and is
@@
 community and begin such dialogs. As many of these database
-researchers are current collaborators of the organizers (see Section
-XXX below for a list of intended invitees), we believe that the
+researchers are current collaborators of the organizers (see Section~\ref{sec:invitees} below for a list of intended invitees), we believe that the
 organizing team is well suited to influence research not only in
@@
 a format has been successfully employed by other workshops in previous
-years at VIS, most notably by the BELIV workshop~\cite{beliv}.
+years at VIS, most notably by the BELIV workshop~\citeref{beliv}.
 
 \section{List of Participants and Invitees}
+\label{sec:invitees}
 
@@
 the industry. If possible, we will also seek to identify a more senior
-person from this invited list as the keynote speaker for the workshop.
+person from this invited list as the keynote speaker for the workshop.\\
 
-Senior researchers:
-
+\noindent \textbf{Senior researchers}:
 \begin{itemize}
+\setlength\itemsep{0em}
 \item Sam Madden (MIT)
@@
 
-Early or mid-career researchers:
+\noindent \textbf{Early or mid-career researchers}:
 \begin{itemize}
+\setlength\itemsep{0em}
 \item Eugene Wu (Columbia)
@@
 
-Junior Researchers: (Students, etc)
+\noindent \textbf{Junior Researchers (Students, etc.)}: 
 \begin{itemize}
+\setlength\itemsep{0em}
 \item Manasi Vartak (MIT, SeeDB)
@@
 
-Relevant companies we will reach out to:
+\noindent \textbf{Relevant companies we will reach out to}:
 \begin{itemize}
+\setlength\itemsep{0em}
 \item Tableau
@@
 TBD.
-
-\section{Related publications by Organizers}
-\cite{Battle:2013:DRO, Lins:2013:NFR, Liu:2013:IRT, Fisher:2012:TMI, Kandel:2012:PIS}
-[6] Battle, Chang, Stonebraker. Dynamic Prefetching of Data Tiles for Interactive Visualization. VLDB, 2015 (Conditionally Accepted).
-
-\section{References}
-
-\cite{Agarwal:2013:BQW}
-
-[7] Stonebraker and Cetintemel. "One Size Fits All": An Idea Whose Time Has Come and Gone. ICDE '05
-[9] Mostak. An overview of MapD. http://map-d.com/docs/mapdwhitepaper. 2014
-[10] Cudre-Mauroux et al. SciDB: A Science-Oriented DBMS. VLDB 2009.
 
@@
 
-\bibliographystyle{abbrv}
-\bibliography{paper}
+\bibliographystyleorgref{plain}
+\bibliographyorgref{orgref}
+
+\bibliographystyleref{plain}
+\bibliographyref{ref}
+
+%\bibliographystyle{abbrv}
+%\bibliography{paper}
 \end{document}

### 2015-05-01 paper.tex
changed to use package enumitem. Began editing text.

@@
 \usepackage{times}
+\usepackage{enumitem}
 \usepackage{multibib}
@@
 architectural changes like columnar databases have enabled faster
-visualization systems, these sorts of innovations have been designed
-less for visualization than for accelerating queries in general. In
+visualization systems, these databases are often not specified optimized for data visualization purposes but for general acceleration of queries. In
 this workshop, we will explore the consequences of pushing the
@@
 the industry. If possible, we will also seek to identify a more senior
-person from this invited list as the keynote speaker for the workshop.\\
+person from this invited list as the keynote speaker for the workshop.
 
-\noindent \textbf{Senior researchers}:
-\begin{itemize}
-\setlength\itemsep{0em}
+\subsection*{Senior researchers}
+\begin{itemize}[topsep=0pt, partopsep=0pt, itemsep=-3pt]
 \item Sam Madden (MIT)
@@
 
-\noindent \textbf{Early or mid-career researchers}:
-\begin{itemize}
-\setlength\itemsep{0em}
+\subsection*{Early or mid-career researchers}
+\begin{itemize}[topsep=0pt, partopsep=0pt, itemsep=-3pt]
 \item Eugene Wu (Columbia)
@@
 
-\noindent \textbf{Junior Researchers (Students, etc.)}: 
-\begin{itemize}
-\setlength\itemsep{0em}
+\subsection*{Junior Researchers (Students, etc.)}
+\begin{itemize}[topsep=0pt, partopsep=0pt, itemsep=-3pt]
 \item Manasi Vartak (MIT, SeeDB)
@@
 
-\noindent \textbf{Relevant companies we will reach out to}:
-\begin{itemize}
-\setlength\itemsep{0em}
+\subsection*{Relevant companies we will reach out to}
+\begin{itemize}[topsep=0pt, partopsep=0pt, itemsep=-3pt]
 \item Tableau

### 2015-05-01 paper.tex
final iteration...

@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Remco Chang \and Danyel Fisher \and Jeffrey Heer \and Carlos Scheidegger}
+\author{Remco Chang\\
+	\scriptsize Tufts University 
+	\and Danyel Fisher\\
+	\scriptsize Microsoft Research 
+	\and Jeffrey Heer\\
+	\scriptsize University of Washington
+ 	\and Carlos Scheidegger\\
+ 	\scriptsize University of Arizona}
 \authorfooter{
@@
 %% the only exception to this rule is the \firstsection command
-The goal of this workshop is to foster innovative work in the backends
-that will support the next generation of interactive data analysis
-tools. We envision such backends to require novel insights in database
-technology, algorithm design, and information visualization techniques.
+The goal of this workshop is to foster innovative work in the backend that will support the next generation of interactive data analysis tools. 
+We envision such backends to require novel insights in database technology, algorithm design, and information visualization techniques.
 
-In the current design of database and analytics systems, interactive
-data analysis concerns are usually an afterthought. Although recent
-architectural changes like columnar databases have enabled faster
-visualization systems, these databases are often not specified optimized for data visualization purposes but for general acceleration of queries. In
-this workshop, we will explore the consequences of pushing the
-concerns of interactive data analysis deep into the ``computational
-fabric'' of our systems: the back-end infrastructure of data storage,
-organization, and retrieval. If we design with the knowledge that
-every analysis is eventually read by a human, bound by perceptual
-limitations and latency constraints, what are the consequences for
-database design, or data analysis and visualization algorithms? Are
-there fundamental tradeoffs? What can we give up?
+In the current design of database and analytics systems, interactive data visualization and analysis concerns are usually an afterthought. 
+Although recent architectural changes like columnar databases have enabled faster visualization systems, these databases are often not specifically optimized for data visualization purposes but for general acceleration of queries.
+In this workshop, we will explore the consequences of pushing the concerns of interactive data visualization and analysis deep into the ``computational fabric'' of our systems: the back-end infrastructure of data storage, organization, and retrieval.
+If we design with the knowledge that every analysis is eventually read by a human, bound by perceptual limitations and latency constraints, what are the consequences for database design, or data analysis and visualization algorithms?
+Are there fundamental tradeoffs? 
+What can we give up?
 
-There have been previous workshops in VIS that have investigated
-issues of large data visualization, such as LDAV (and LDAV will be a
-co-located symposium this year at VIS). Our focus is markedly
-different from that of LDAV. LDAV focuses on simulation and imaging
-data, while we believe current techniques in information visualization
-and visual analytics are close to hitting a ceiling in terms of
-performance. We propose this workshop as a venue to discuss these
-limits, and fundamental new ways to attack them.
+There have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV (and LDAV will be a co-located symposium this year at VIS). 
+Our focus is markedly different from that of LDAV. 
+While LDAV focuses on simulation and imaging data, we believe current backend techniques in information visualization and visual analytics are close to hitting a ceiling in terms of performance. 
+We propose this workshop as a venue to discuss these limits, and fundamental new ways to attack them.
 
 \section{Technical Scope}
-The scope of this workshop will span, among possibly other fields,
-information visualization, visual analytics, algorithm design,
-database architecture, and statistics. Although this is clearly a
-broad scope, the workshop will focus on the backends for interactive
-tools. We have made significant progress in our understanding of
-perceptual and human-interaction issues behind interactive
-visualization design. However, much of this understanding has not, in
-general, been translated into new designs for our backend
-infrastructure. We think now is the right time to have a discussion of
-backend design for interactive data analysis. Following the principle
-of ``No One Size Fits All'' in designing databases~\citeref{stonebraker2005one}, recent
-publications by the organizers have shown that by taking advantage of
-needs and limitations of interactive visualization systems, databases
-can be optimized by making tradeoffs in novel and interesting
-ways. For example, imMens~\citeorgref{Liu:2013:IRT}, profiler~\citeorgref{Kandel:2012:PIS}, and nanocubes~\citeorgref{Lins:2013:NFR} have
-shown that careful algorithm design can enable extremely fast
-interactive visualizations of massive data. Scalr~\citeorgref{Battle:2013:DRO} has shown that
-traditional query planners can fruitfully take advantage of perceptual
-and resolution limits. ForeCache~\citeorgref{Battle:2015:VLDB} utilizes a predictive prefetching
-strategy by learning a user’s interaction patterns and prefetching
-data from server to support real-time visual exploration. Lastly,
-SampleAction~\citeorgref{Fisher:2012:TMI} takes advantage of streaming databases and
-incorporates HCI principles that users need immediate feedback, but
-are willing to wait for high-fidelity results.
+The scope of this workshop will span, among possibly other fields, information visualization, visual analytics, algorithm design, database architecture, and statistics. 
+Although this is clearly a broad scope, the workshop will focus on backends for interactive tools. 
+As a community, visualization researchers have made significant progress in our understanding of perceptual and human-interaction issues behind interactive visualization design. 
+However, much of this understanding has not, in general, been translated into new designs for our backend infrastructure. 
+We think now is the right time to have a discussion of backend design for interactive data analysis. 
+Following the principle of ``No One Size Fits All'' in designing databases~\citeref{stonebraker2005one}, recent publications by the organizers have shown that by taking advantage of the needs and limitations of interactive visualization systems, databases can be optimized by making tradeoffs in novel and interesting
+ways. 
+For example, imMens~\citeorgref{Liu:2013:IRT}, profiler~\citeorgref{Kandel:2012:PIS}, and nanocubes~\citeorgref{Lins:2013:NFR} have shown that careful algorithm design can enable extremely fast interactive visualizations of massive data.
+Scalr~\citeorgref{Battle:2013:DRO} has shown that traditional query planners can fruitfully take advantage of perceptual and resolution limits. 
+ForeCache~\citeorgref{Battle:2015:VLDB} utilizes a predictive prefetching strategy by learning a user's interaction patterns and prefetching data from the server to support real-time visual data exploration.
+Lastly, SampleAction~\citeorgref{Fisher:2012:TMI} takes advantage of streaming databases and
+incorporates the HCI principle that users need immediate feedback, but are willing to wait for high-fidelity results.
 
-In addition to the work by the organizers, researchers in the database
-community have also made significant stride in developing databases
-specifically for interactive visualization, analysis, and exploration
-of data. For example, systems like Tableau utilizes columnar databases
-which provide speedups for typical visual analytics tasks that require
-operations across columns (instead of the traditional rows). BlinkDB~\citeref{Agarwal:2013:BQW}
-uses a sampling strategy to build a smaller, but representative
-database from a larger one. Map-D~\citeref{mapd} takes advantage of modern
-graphics card capabilities and resulted in a specialized database that
-runs entirely within (an array of) graphics cards. Finally, SciDB~\citeref{cudre2009demonstration}
-breaks away from the traditional relational data structure and is
-designed to store and query multidimensional scientific data that are
-stored in an array-based structure.
+In addition to the work by the organizers, researchers in the database community have also made significant strides in developing databases specifically for interactive visualization, analysis, and exploration
+of data. 
+For example, systems like Tableau~\citeref{stolte2002polaris} utilize columnar databases which provide speedups for typical visual analytics tasks that require operations across columns (instead of the traditional rows).
+BlinkDB~\citeref{Agarwal:2013:BQW} uses a sampling strategy to build a smaller, but representative database from a larger one. 
+MapD~\citeref{mapd} takes advantage of modern graphics card capabilities and resulted in a specialized database that runs entirely within (an array of) graphics cards. 
+Finally, SciDB~\citeref{cudre2009demonstration} breaks away from the traditional relational data structure and is designed to efficiently store and query multidimensional scientific data that are stored in an array-based structure.
 
-Given the increasing number of publications and backend systems that
-can be tailored or applied to interactive visual exploration of large
-data, we believe that it is very important for the InfoVis and the
-VAST communities to begin having discussions and to develop a research
-agenda that focuses on the backend technologies. In this proposed
-workshop, we seek to invite members from the database community
-(including authors of the aforementioned papers) to meet with the VIS
-community and begin such dialogs. As many of these database
-researchers are current collaborators of the organizers (see Section~\ref{sec:invitees} below for a list of intended invitees), we believe that the
-organizing team is well suited to influence research not only in
-information visualization, but in these other related fields as well.
+Given the increasing number of publications and backend systems that can be tailored or applied to interactive visual exploration of large data, we believe that it is very important for the InfoVis and the
+VAST communities to begin having discussions and to develop a research agenda that focuses on backend technologies. 
+In this proposed workshop, we seek to invite members from the database community (including authors of the aforementioned papers) to meet with the VIS community and begin such dialogs. 
+As many of these database researchers are current collaborators of the organizers (see Section~\ref{sec:invitees} below for a list of intended invitees), we believe that the organizing team is well suited to influence research not only in information visualization, but in these other related fields as well.
 
 \section{Format, Length, and Planned Activities}
-We propose a half-day workshop that focuses on discussion between the
-VIS and database researchers. The workshop will solicit submissions of
-position papers as well as short research papers from both the VIS and
-database communities. These papers will be reviewed by the organizing
-team. The authors of the selected top papers will be asked to each
-give a short presentation of their work (or position).
+We propose a half-day workshop that focuses on generating discussions between VIS and database researchers. 
+The workshop will solicit submissions of position papers as well as short research papers from both the VIS and database communities. 
+These papers will be reviewed by the organizing team. 
+The authors of the selected top papers will be invited to each give a short presentation of their work (or position).
 
-The goal of the presentation is to allow the attendees to get an
-overview of the different perspectives and approaches of data storage,
-organization, and retrieval techniques that can be tailored
-specifically to support interactive data visualization and
-exploration. We anticipate that these presentations will take up
-approximately the first half of the workshop.
+The goal of the presentation is to allow the attendees to gain an overview of the different perspectives and approaches of data storage, organization, and retrieval techniques that can be tailored specifically to support interactive data visualization and exploration.
+We anticipate that these presentations will take up approximately the first half of the workshop.
 
-Following a short break, the second half of the workshop will be
-oriented around open discussions and possibly a question-and-answer
-session. As we anticipate that members of the two different
-communities will not have a common language or viewpoint, we believe
-that the open-ended format will facilitate discussions and general
-knowledge sharing.
+Following a short break, the second half of the workshop will be oriented around open discussions and possibly a question-and-answer session. 
+As we anticipate that members of the two different communities will not have a common language or viewpoint, we believe that the open-ended format will facilitate discussions and general knowledge sharing.
 
-It is relevant to note that the organizing team is also currently
-seeking a followup ``closed-door'' (invitation only) meeting that will
-take place in the afternoon outside of the main VIS venue. The purpose
-of this closed-door meeting is to allow for a more intimate discussion
-between enthusiastic researchers who are actively seeking
-collaborations across the aisle. Although it is not the goal of the
-organizing team to be exclusive, it is our belief that a more focused
-discussion will have a higher likelihood towards developing a cohesive
-research agenda. Note that this is not an extraordinary proposal: such
-a format has been successfully employed by other workshops in previous
-years at VIS, most notably by the BELIV workshop~\citeref{beliv}.
+It is relevant to note that the organizing team is also currently seeking a follow-up ``closed-door'' (invitation only) meeting that will take place in the afternoon outside of the main VIS venue. 
+The purpose of this closed-door meeting is to allow for a more intimate discussion between enthusiastic researchers who are actively seeking collaborations across the aisle. 
+Although it is not the goal of the organizing team to be exclusive, it is our belief that a more focused
+discussion will have a higher likelihood towards developing a cohesive research agenda. 
+Note that this is not an extraordinary proposal: such an off-site format has been successfully employed by other workshops in previous years at VIS, most notably by the BELIV workshop~\citeref{beliv}.
 
@@
 
-As noted earlier, it is our hope that this workshop will include both
-attendees from the VIS and database communities. However, since the
-workshop will be located at VIS, we will specifically reach out to
-members of the database community and invite them to attend the
-workshop. We will seek a mix of researchers in different stages of
-their careers, together with researchers from different companies in
-the industry. If possible, we will also seek to identify a more senior
-person from this invited list as the keynote speaker for the workshop.
+As noted earlier, it is our hope that this workshop will include both attendees from the VIS and database communities. 
+However, since the workshop will be located at VIS, we will specifically reach out to members of the database community and invite them to attend the workshop. 
+We will seek a mix of researchers in different stages of their careers, together with researchers from different companies in the industry. 
+If possible, we will also seek to identify a more senior person from this invited list as the keynote speaker for the workshop.
 
@@
 \item Arnab Nandi (OSU)
-\item Cecilia Aragon (UW; astronomy/scivis background)
+\item Cecilia Aragon (UW)
 \item Jenny Duggan (Northwestern)
@@
 
-\subsection*{Junior Researchers (Students, etc.)}
+\subsection*{Junior researchers (students, etc.)}
 \begin{itemize}[topsep=0pt, partopsep=0pt, itemsep=-3pt]
-\item Manasi Vartak (MIT, SeeDB)
+\item Manasi Vartak (MIT)
 \item Todd Mostak (MapD)
-\item Leilani Battle (MIT, Tile Prefetching)
+\item Leilani Battle (MIT)
 \item Dominik Moritz (UW)
-\item Holger Pirk (MIT postdoc, databases)
+\item Holger Pirk (MIT postdoc)
 \end{itemize}
@@
 
-We should ask for a setup split halfway with seats for lectures and space for posters in the back. TBD.
+In order to facilitate both paper presentations and open discussions, we seek to have a room that has space for both seats and for posters.
+A podium with audio/video support will be necessary for the paper presentations.
+In addition, some supplies for sketching and drawing would be beneficial in facilitating group discussions (e.g. big Easel Pads with markers).
 
@@
 
-TBD.
+There are three goals for this workshop. 
+First, we hope to engage both researchers in the VIS and the database communities.
+Should we be successful in attracting and inviting database researchers, the workshop can serve as a forum for true interdisciplinary discussion and collaboration.
+
+Second, as we will accept both position papers and (short) research papers, we aim to compile the accepted papers into a short proceeding that we will disseminate to the workshop attendees. 
+While we do not anticipate this proceeding to contain cutting-edge research contributions, we do envision the content to serve as a starting point for an open dialog between the communities.
+
+Lastly, depending on the outcomes of these discussions, we aim to sketch out an initial outline for a research agenda for this burgeoning field.
+This research agenda will summarize the current state-of-the-art in both the VIS and database communities concerning backend supports for interactive data visualization and analysis.
+In addition, it will highlight the top challenges that need to be addressed that will hopefully spur future research in this field.
+The four organizers will enlist attendees of the workshop to help author this research agenda with the hope that the resulting paper will be published in a venue that is visible to both the VIS and database communities.
 

### 2017-02-09 2017/paper.tex
setting up 2017 submission
edits...

--- 2015-05-01 paper.tex
@@
 
-\title{TRIAD: Towards Real-time, Interactive Analysis of Data}
+\title{DSIA: Data Systems for Interactive Analysis}
 
@@
 %% ``\begin{document}'' command. It prepares and prints the title block.
+\section{Background: DSIA 2015}
+The primary goal of the DSIA workshop is to bring together researchers from the Database community to participate in VIS.  
+The premise of the workshop is that the development of back-end data systems is of increasing importance for visualization tools due to the growing size and complexity of data and the increased demand of interactivity.
+However, tackling the issue of data systems cannot be addressed by VIS researchers alone. 
+Collaboration with other communities is essential to ensure integration between the front-end visualization and back-end storage and computation.
+
+The first DSIA workshop was held in VIS 2015 with somewhere around 150 to 200 attendees. 
+We were able to bring in young, but prominent researchers in Databases, including Eugene Wu (Columbia), Arnab Nandi (Ohio State), Aditya Parameswaran (Illinois), Aaron Elmore (Univ of Chicago), Jenny Duggan (Northwestern), Leilani Battle (MIT); as well as senior researchers like Marti Hearst (Berkeley) and Joe Hellerstein (Berkeley, who was also the invited keynote speaker).
+Overall, the event was successful -- we observed the researchers from both sides interacting with each other, and anecdotally we are aware of a few collaborative projects borne from these discussions.
+For example, the collaboration between Jean-Daniel Fekete and Tim Kraska (Brown Univ) resulted in a recent TVCG paper.
+
 \section{Goals}

### 2017-02-10 2017/paper.tex
clean up...mostly ready to go. Need to specify dates

@@
 Overall, the event was successful -- we observed the researchers from both sides interacting with each other, and anecdotally we are aware of a few collaborative projects borne from these discussions.
-For example, the collaboration between Jean-Daniel Fekete and Tim Kraska (Brown Univ) resulted in a recent TVCG paper.
+For example, the collaboration between Jean-Daniel Fekete and Tim Kraska (Brown University) resulted in a recent TVCG paper~\citeref{zgraggen2016progressive}.
 
@@
 %% the only exception to this rule is the \firstsection command
-The goal of this workshop is to foster innovative work in the backend that will support the next generation of interactive data analysis tools. 
+The goal of the 2017 workshop is to continue the community-building activities that started in 2015, which is to foster innovative work in the backend that will support the next generation of interactive data analysis tools.
 We envision such backends to require novel insights in database technology, algorithm design, and information visualization techniques.
@@
 Although recent architectural changes like columnar databases have enabled faster visualization systems, these databases are often not specifically optimized for data visualization purposes but for general acceleration of queries.
-In this workshop, we will explore the consequences of pushing the concerns of interactive data visualization and analysis deep into the ``computational fabric'' of our systems: the back-end infrastructure of data storage, organization, and retrieval.
+In this workshop, we will explore the consequences of pushing the concerns of interactive data visualization and analysis deep into the ``computational fabric'' of our systems: the backend infrastructure of data storage, organization, and retrieval.
 If we design with the knowledge that every analysis is eventually read by a human, bound by perceptual limitations and latency constraints, what are the consequences for database design, or data analysis and visualization algorithms?
@@
 
-There have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV (and LDAV will be a co-located symposium this year at VIS). 
+Beyond DSIA in 2015, here have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV. 
 Our focus is markedly different from that of LDAV. 
@@
 We propose this workshop as a venue to discuss these limits, and fundamental new ways to attack them.
+
+The workshop most similar to DSIA is the HILDA workshop, which was held at SIGMOD in 2016 (and will be held again in 2017 in SIGMOD). 
+The key difference between the two is that the database researchers at HILDA approach the problem from the opposite end as DSIA -- in DSIA we seek to leverage the user's abilities (and limitations) in designing backend systems, in HILDA the researchers start with the backend system and analyze how the system might be useful to the user in data analysis tasks.
 
@@
 Scalr~\citeorgref{Battle:2013:DRO} has shown that traditional query planners can fruitfully take advantage of perceptual and resolution limits. 
-ForeCache~\citeorgref{Battle:2015:VLDB} utilizes a predictive prefetching strategy by learning a user's interaction patterns and prefetching data from the server to support real-time visual data exploration.
+ForeCache~\citeorgref{battle2016dynamic} utilizes a predictive prefetching strategy by learning a user's interaction patterns~\citeorgref{brown2014finding} and prefetching data from the server to support real-time visual data exploration.
 Lastly, SampleAction~\citeorgref{Fisher:2012:TMI} takes advantage of streaming databases and
@@
 The goal of the presentation is to allow the attendees to gain an overview of the different perspectives and approaches of data storage, organization, and retrieval techniques that can be tailored specifically to support interactive data visualization and exploration.
-We anticipate that these presentations will take up approximately the first half of the workshop.
+We anticipate that these presentations will take up approximately two thirds of the workshop.
 
-Following a short break, the second half of the workshop will be oriented around open discussions and possibly a question-and-answer session. 
+Following a short break, the last one third of the workshop will be oriented around open discussions and possibly a question-and-answer session. 
 As we anticipate that members of the two different communities will not have a common language or viewpoint, we believe that the open-ended format will facilitate discussions and general knowledge sharing.
+Based on our experience in 2015, a round-table discussion involving 150 to 200 participants is not possible.
+As a result, we envision a panel style discussion where the database researchers sit on stage and the participants from the VIS community can ask them questions and initiate a dialog.
 
@@
 \item Joe Hellerstein (Berkeley, Trifacta)
+\item Marti Hearst (Berkeley)
 \item Mike Stonebraker (MIT)
@@
 \item Stan Zdonik (Brown)
+\item Juliana Freire (NYU)
 \end{itemize}
@@
 \item Eugene Wu (Columbia)
-\item Aaron Elmore (U. Chicago)
+\item Aaron Elmore (Chicago)
+\item Tim Kraska (Brown)
+\item Carsten Binning (Brown)
+\item Aditya Parameswaran (UIUC)
+\item Arnab Nandi (Ohio State)
 \item Joey Gonzalez (Berkeley AMPLab, CMU, GraphLab/Dato)
@@
 \item Dominik Moritz (UW)
-\item Holger Pirk (MIT postdoc)
+%\item Holger Pirk (MIT postdoc)
 \end{itemize}
@@
 \item Graphistry
+\item Autodesk
 \end{itemize}
@@
 
+\section{Important Dates}
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
+XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
+
+Need to specify 3 dates:
+(1) date of participation (when the workshop will be held); (2) submission deadline (for the workshop), (3) date by which we'll notice by the authors
+
 \section{Organizer Bios}
 
-\paragraph*{Remco Chang} is an Assistant Professor in the Computer Science Department at Tufts University.
+\paragraph*{Remco Chang} is an Associate Professor in the Computer Science Department at Tufts University.
 He received his BS from Johns Hopkins University in 1997 in Computer Science and Economics, MSc from Brown University in 2000, and PhD in computer science from UNC Charlotte in 2009.
 Prior to his PhD, he worked for Boeing developing real-time flight tracking and visualization software, followed by a position at UNC Charlotte as a research scientist.
-His current research interests include visual analytics, information visualization, and HCI.
+His current research interests include visual analytics, information visualization, HCI, and databases.
 His research has been funded by NSF, DHS, MIT Lincoln Lab, and Draper.

### 2017-02-22 2017/paper.tex
dates

@@
 \section{Important Dates}
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
-XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
 
-Need to specify 3 dates:
-(1) date of participation (when the workshop will be held); (2) submission deadline (for the workshop), (3) date by which we'll notice by the authors
+These dates are tentative, because we want to align them to the early
+registration deadline of IEEE VIS 2017. This date has not been
+announced yet, and we will adjust our timeline accordingly. We are
+assuming the early registration deadline is 6 weeks before the
+conference: roughly August 15th.
+
+\begin{itemize}
+  \item Date of participation: Sunday, Monday, or Tuesday (Oct 1st,
+    2nd, or 3rd)
+  \item Submission Deadline: July 15th.
+  \item Notification Deadline: August 1st.
+\end{itemize}
 

### 2017-02-22 2017/paper.tex
to hell with your fancy citeref citeorgref or whatever. it wouldn't build. now with plain cite. it build. ship it

@@
 Overall, the event was successful -- we observed the researchers from both sides interacting with each other, and anecdotally we are aware of a few collaborative projects borne from these discussions.
-For example, the collaboration between Jean-Daniel Fekete and Tim Kraska (Brown University) resulted in a recent TVCG paper~\citeref{zgraggen2016progressive}.
+For example, the collaboration between Jean-Daniel Fekete and Tim Kraska (Brown University) resulted in a recent TVCG paper~\cite{zgraggen2016progressive}.
 
@@
 We think now is the right time to have a discussion of backend design for interactive data analysis. 
-Following the principle of ``No One Size Fits All'' in designing databases~\citeref{stonebraker2005one}, recent publications by the organizers have shown that by taking advantage of the needs and limitations of interactive visualization systems, databases can be optimized by making tradeoffs in novel and interesting
+Following the principle of ``No One Size Fits All'' in designing databases~\cite{stonebraker2005one}, recent publications by the organizers have shown that by taking advantage of the needs and limitations of interactive visualization systems, databases can be optimized by making tradeoffs in novel and interesting
 ways. 
-For example, imMens~\citeorgref{Liu:2013:IRT}, profiler~\citeorgref{Kandel:2012:PIS}, and nanocubes~\citeorgref{Lins:2013:NFR} have shown that careful algorithm design can enable extremely fast interactive visualizations of massive data.
-Scalr~\citeorgref{Battle:2013:DRO} has shown that traditional query planners can fruitfully take advantage of perceptual and resolution limits. 
-ForeCache~\citeorgref{battle2016dynamic} utilizes a predictive prefetching strategy by learning a user's interaction patterns~\citeorgref{brown2014finding} and prefetching data from the server to support real-time visual data exploration.
-Lastly, SampleAction~\citeorgref{Fisher:2012:TMI} takes advantage of streaming databases and
+For example, imMens~\cite{Liu:2013:IRT}, profiler~\cite{Kandel:2012:PIS}, and nanocubes~\cite{Lins:2013:NFR} have shown that careful algorithm design can enable extremely fast interactive visualizations of massive data.
+Scalr~\cite{Battle:2013:DRO} has shown that traditional query planners can fruitfully take advantage of perceptual and resolution limits. 
+ForeCache~\cite{battle2016dynamic} utilizes a predictive prefetching strategy by learning a user's interaction patterns~\cite{brown2014finding} and prefetching data from the server to support real-time visual data exploration.
+Lastly, SampleAction~\cite{Fisher:2012:TMI} takes advantage of streaming databases and
 incorporates the HCI principle that users need immediate feedback, but are willing to wait for high-fidelity results.
@@
 of data. 
-For example, systems like Tableau~\citeref{stolte2002polaris} utilize columnar databases which provide speedups for typical visual analytics tasks that require operations across columns (instead of the traditional rows).
-BlinkDB~\citeref{Agarwal:2013:BQW} uses a sampling strategy to build a smaller, but representative database from a larger one. 
-MapD~\citeref{mapd} takes advantage of modern graphics card capabilities and resulted in a specialized database that runs entirely within (an array of) graphics cards. 
-Finally, SciDB~\citeref{cudre2009demonstration} breaks away from the traditional relational data structure and is designed to efficiently store and query multidimensional scientific data that are stored in an array-based structure.
+For example, systems like Tableau~\cite{stolte2002polaris} utilize columnar databases which provide speedups for typical visual analytics tasks that require operations across columns (instead of the traditional rows).
+BlinkDB~\cite{Agarwal:2013:BQW} uses a sampling strategy to build a smaller, but representative database from a larger one. 
+MapD~\cite{mapd} takes advantage of modern graphics card capabilities and resulted in a specialized database that runs entirely within (an array of) graphics cards. 
+Finally, SciDB~\cite{cudre2009demonstration} breaks away from the traditional relational data structure and is designed to efficiently store and query multidimensional scientific data that are stored in an array-based structure.
 
@@
 discussion will have a higher likelihood towards developing a cohesive research agenda. 
-Note that this is not an extraordinary proposal: such an off-site format has been successfully employed by other workshops in previous years at VIS, most notably by the BELIV workshop~\citeref{beliv}.
+Note that this is not an extraordinary proposal: such an off-site format has been successfully employed by other workshops in previous years at VIS, most notably by the BELIV workshop~\cite{beliv}.
 
@@
 
-\bibliographystyleorgref{plain}
-\bibliographyorgref{orgref}
+%% \bibliographystyleorgref{plain}
+%% \bibliographyorgref{orgref}
 
-\bibliographystyleref{plain}
-\bibliographyref{ref}
+%% \bibliographystyleref{plain}
+%% \bibliographyref{ref}
 
-%\bibliographystyle{abbrv}
-%\bibliography{paper}
+%% \bibliographystyleref{plain}
+%% \bibliographyref{orgref}
+\bibliographystyle{plain}
+\bibliography{paper}
+
 \end{document}

### 2017-02-22 2017/paper.tex
tiny typo

@@
 The first DSIA workshop was held in VIS 2015 with somewhere around 150 to 200 attendees. 
-We were able to bring in young, but prominent researchers in Databases, including Eugene Wu (Columbia), Arnab Nandi (Ohio State), Aditya Parameswaran (Illinois), Aaron Elmore (Univ of Chicago), Jenny Duggan (Northwestern), Leilani Battle (MIT); as well as senior researchers like Marti Hearst (Berkeley) and Joe Hellerstein (Berkeley, who was also the invited keynote speaker).
+We were able to bring in young, but prominent researchers in Databases, including Eugene Wu (Columbia), Arnab Nandi (Ohio State), Aditya Parameswaran (Illinois), Aaron Elmore (Univ.
+of Chicago), Jenny Duggan (Northwestern), Leilani Battle (MIT); as well as senior researchers like Marti Hearst (Berkeley) and Joe Hellerstein (Berkeley, who was also the invited keynote speaker).
 Overall, the event was successful -- we observed the researchers from both sides interacting with each other, and anecdotally we are aware of a few collaborative projects borne from these discussions.

### 2017-02-23 2017/paper.tex
last revision of dsia 2017

@@
 
-\title{DSIA: Data Systems for Interactive Analysis}
+\title{DSIA 2017: Data Systems for Interactive Analysis}
 
@@
 
-Beyond DSIA in 2015, here have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV. 
+Beyond DSIA in 2015, there have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV. 
 Our focus is markedly different from that of LDAV. 
@@
 \section{Format, Length, and Planned Activities}
-We propose a half-day workshop that focuses on generating discussions between VIS and database researchers. 
+We propose a \textbf{half-day workshop} that focuses on generating discussions between VIS and database researchers. 
 The workshop will solicit submissions of position papers as well as short research papers from both the VIS and database communities. 
@@
 The authors of the selected top papers will be invited to each give a short presentation of their work (or position).
+
+\begin{center}
+\begin{tabular}{ l l } 
+ \hline
+ Schedule & Event \\ 
+ \hline
+ 8:30  & Welcome and Introduction\\ 
+ 8:45 & Invited Keynote Speaker\\ 
+ 9:30 & Paper Session 1\\ 
+ 10:10 & Coffee Break \\ 
+ 10:30  & Paper Session 2\\ 
+ 11:30  & Panel Discussion\\ 
+ 12:00 & Closing\\
+ 12:10 & Lunch Break\\
+ \hline
+\end{tabular}
+\end{center}
 
@@
 
-\begin{itemize}
-  \item Date of participation: Sunday, Monday, or Tuesday (Oct 1st,
-    2nd, or 3rd)
-  \item Submission Deadline: July 15th.
-  \item Notification Deadline: August 1st.
-\end{itemize}
+\begin{center}
+\begin{tabular}{ l l } 
+ \hline
+ Date & Event \\ 
+ \hline
+  Date of Participation & Sunday (Oct 1st), or \\
+					& Monday (Oct 2nd), or \\
+					& Tuesday (Oct 3rd)\\
+  Submission Deadline & July 15th \\
+  Notification Date & August 1st\\
+  Camera-Ready Deadline & August 15th\\
+  USB Proceeding Deadline & August 25th\\
+ \hline
+\end{tabular}
+\end{center}
+
+%\begin{itemize}
+%  \item Date of participation: Sunday, Monday, or Tuesday (Oct 1st,
+%    2nd, or 3rd)
+%  \item Submission Deadline: July 15th.
+%  \item Notification Deadline: August 1st.
+%  \item 
+%\end{itemize}
 
@@
 His current research interests include visual analytics, information visualization, HCI, and databases.
-His research has been funded by NSF, DHS, MIT Lincoln Lab, and Draper.
+His research has been funded by NSF, DARPA, DHS, MIT Lincoln Lab, and Draper.
 He received the NSF CAREER Award in 2015.

### 2019-07-09 2015/paper.tex
add

--- 2017-02-22 2017/paper.tex
@@
 \usepackage{enumitem}
-\usepackage{multibib}
-\newcites{orgref}{Publications by Organizers}
-\newcites{ref}{Other References}
+%\usepackage{multibib}
+%\newcites{orgref}{Publications by Organizers}
+%\newcites{ref}{Other References}
 
@@
 
-\title{DSIA: Data Systems for Interactive Analysis}
+\title{TRIAD: Towards Real-time, Interactive Analysis of Data}
 
@@
 %% ``\begin{document}'' command. It prepares and prints the title block.
-\section{Background: DSIA 2015}
-The primary goal of the DSIA workshop is to bring together researchers from the Database community to participate in VIS.  
-The premise of the workshop is that the development of back-end data systems is of increasing importance for visualization tools due to the growing size and complexity of data and the increased demand of interactivity.
-However, tackling the issue of data systems cannot be addressed by VIS researchers alone. 
-Collaboration with other communities is essential to ensure integration between the front-end visualization and back-end storage and computation.
-
-The first DSIA workshop was held in VIS 2015 with somewhere around 150 to 200 attendees. 
-We were able to bring in young, but prominent researchers in Databases, including Eugene Wu (Columbia), Arnab Nandi (Ohio State), Aditya Parameswaran (Illinois), Aaron Elmore (Univ.
-of Chicago), Jenny Duggan (Northwestern), Leilani Battle (MIT); as well as senior researchers like Marti Hearst (Berkeley) and Joe Hellerstein (Berkeley, who was also the invited keynote speaker).
-Overall, the event was successful -- we observed the researchers from both sides interacting with each other, and anecdotally we are aware of a few collaborative projects borne from these discussions.
-For example, the collaboration between Jean-Daniel Fekete and Tim Kraska (Brown University) resulted in a recent TVCG paper~\cite{zgraggen2016progressive}.
-
 \section{Goals}
 %% the only exception to this rule is the \firstsection command
-The goal of the 2017 workshop is to continue the community-building activities that started in 2015, which is to foster innovative work in the backend that will support the next generation of interactive data analysis tools.
+The goal of this workshop is to foster innovative work in the backend that will support the next generation of interactive data analysis tools. 
 We envision such backends to require novel insights in database technology, algorithm design, and information visualization techniques.
@@
 Although recent architectural changes like columnar databases have enabled faster visualization systems, these databases are often not specifically optimized for data visualization purposes but for general acceleration of queries.
-In this workshop, we will explore the consequences of pushing the concerns of interactive data visualization and analysis deep into the ``computational fabric'' of our systems: the backend infrastructure of data storage, organization, and retrieval.
+In this workshop, we will explore the consequences of pushing the concerns of interactive data visualization and analysis deep into the ``computational fabric'' of our systems: the back-end infrastructure of data storage, organization, and retrieval.
 If we design with the knowledge that every analysis is eventually read by a human, bound by perceptual limitations and latency constraints, what are the consequences for database design, or data analysis and visualization algorithms?
@@
 
-Beyond DSIA in 2015, here have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV. 
+There have been previous workshops in VIS that have investigated issues of large data visualization, such as LDAV (and LDAV will be a co-located symposium this year at VIS). 
 Our focus is markedly different from that of LDAV. 
@@
 We propose this workshop as a venue to discuss these limits, and fundamental new ways to attack them.
-
-The workshop most similar to DSIA is the HILDA workshop, which was held at SIGMOD in 2016 (and will be held again in 2017 in SIGMOD). 
-The key difference between the two is that the database researchers at HILDA approach the problem from the opposite end as DSIA -- in DSIA we seek to leverage the user's abilities (and limitations) in designing backend systems, in HILDA the researchers start with the backend system and analyze how the system might be useful to the user in data analysis tasks.
 
@@
 Scalr~\cite{Battle:2013:DRO} has shown that traditional query planners can fruitfully take advantage of perceptual and resolution limits. 
-ForeCache~\cite{battle2016dynamic} utilizes a predictive prefetching strategy by learning a user's interaction patterns~\cite{brown2014finding} and prefetching data from the server to support real-time visual data exploration.
+ForeCache~\cite{Battle:2015:VLDB} utilizes a predictive prefetching strategy by learning a user's interaction patterns and prefetching data from the server to support real-time visual data exploration.
 Lastly, SampleAction~\cite{Fisher:2012:TMI} takes advantage of streaming databases and
@@
 The goal of the presentation is to allow the attendees to gain an overview of the different perspectives and approaches of data storage, organization, and retrieval techniques that can be tailored specifically to support interactive data visualization and exploration.
-We anticipate that these presentations will take up approximately two thirds of the workshop.
+We anticipate that these presentations will take up approximately the first half of the workshop.
 
-Following a short break, the last one third of the workshop will be oriented around open discussions and possibly a question-and-answer session. 
+Following a short break, the second half of the workshop will be oriented around open discussions and possibly a question-and-answer session. 
 As we anticipate that members of the two different communities will not have a common language or viewpoint, we believe that the open-ended format will facilitate discussions and general knowledge sharing.
-Based on our experience in 2015, a round-table discussion involving 150 to 200 participants is not possible.
-As a result, we envision a panel style discussion where the database researchers sit on stage and the participants from the VIS community can ask them questions and initiate a dialog.
 
@@
 \item Joe Hellerstein (Berkeley, Trifacta)
-\item Marti Hearst (Berkeley)
 \item Mike Stonebraker (MIT)
@@
 \item Stan Zdonik (Brown)
-\item Juliana Freire (NYU)
 \end{itemize}
@@
 \item Eugene Wu (Columbia)
-\item Aaron Elmore (Chicago)
-\item Tim Kraska (Brown)
-\item Carsten Binning (Brown)
-\item Aditya Parameswaran (UIUC)
-\item Arnab Nandi (Ohio State)
+\item Aaron Elmore (U. Chicago)
 \item Joey Gonzalez (Berkeley AMPLab, CMU, GraphLab/Dato)
@@
 \item Dominik Moritz (UW)
-%\item Holger Pirk (MIT postdoc)
+\item Holger Pirk (MIT postdoc)
 \end{itemize}
@@
 \item Graphistry
-\item Autodesk
 \end{itemize}
@@
 
-\section{Important Dates}
-
-These dates are tentative, because we want to align them to the early
-registration deadline of IEEE VIS 2017. This date has not been
-announced yet, and we will adjust our timeline accordingly. We are
-assuming the early registration deadline is 6 weeks before the
-conference: roughly August 15th.
-
-\begin{itemize}
-  \item Date of participation: Sunday, Monday, or Tuesday (Oct 1st,
-    2nd, or 3rd)
-  \item Submission Deadline: July 15th.
-  \item Notification Deadline: August 1st.
-\end{itemize}
-
 \section{Organizer Bios}
 
-\paragraph*{Remco Chang} is an Associate Professor in the Computer Science Department at Tufts University.
+\paragraph*{Remco Chang} is an Assistant Professor in the Computer Science Department at Tufts University.
 He received his BS from Johns Hopkins University in 1997 in Computer Science and Economics, MSc from Brown University in 2000, and PhD in computer science from UNC Charlotte in 2009.
 Prior to his PhD, he worked for Boeing developing real-time flight tracking and visualization software, followed by a position at UNC Charlotte as a research scientist.
-His current research interests include visual analytics, information visualization, HCI, and databases.
+His current research interests include visual analytics, information visualization, and HCI.
 His research has been funded by NSF, DHS, MIT Lincoln Lab, and Draper.
@@
 
-%% \bibliographystyleorgref{plain}
-%% \bibliographyorgref{orgref}
+%\bibliographystyleorgref{plain}
+%\bibliographyorgref{orgref}
 
-%% \bibliographystyleref{plain}
-%% \bibliographyref{ref}
+%\bibliographystyleref{plain}
+%\bibliographyref{ref}
 
-%% \bibliographystyleref{plain}
-%% \bibliographyref{orgref}
-\bibliographystyle{plain}
-\bibliography{paper}
-
+\bibliographystyle{abbrv}
+\bibliography{ref}
 \end{document}
